---
title: "National Parks API Vignette"
author: "Kelley Breeze"
date: "2022-09-22"
output: 
  html_document: 
    toc: yes 
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Requirements

```{r}
library(jsonlite)
library(tidyverse)
```


# National Parks Service API


This function allows user to access the `campgrounds` endpoint on the National Parks Service API. They can specify two arguments, one for `stateAbbreviation` and a second for `limitResultsTo`. If no information is entered for these, the default values are "NC" and "100".

```{r}
getCampgrounds<- function(stateAbbreviation="NC", limitResultsTo="100"){
  baseURL<-"https://developer.nps.gov/api/v1/campgrounds?"
  
  state<- paste("stateCode=", stateAbbreviation=stateAbbreviation, sep = "")
  
  limit<- paste("limit=", limitResultsTo=limitResultsTo, sep = "")
  
  searchURL<- paste(baseURL,state,"&", limit, "&", apiKey, sep = "")
  
  campgrounds<- fromJSON(searchURL)
  return(campgrounds$data)

}

```



Test `getCampgrounds()` function using default settings
```{r}
defaultCampResults<- getCampgrounds()

defaultCampResults
```



Test `getCampgrounds()` function with user specifications of `stateAbbreviation` of `CA` and `limitResultsTo` of `30`.

```{r}
testCampResultsCA30<-getCampgrounds(stateAbbreviation = "CA", limitResultsTo = "30")

testCampResultsCA30
```

